A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker

### 2019-08-20 rr-flow.Rmd
added ideas on checksums added ideas on licensing in summary

---
title: "A Reproducible Workflow with RMarkdown, Git, Make & Docker"
shorttitle: "Reproducible Workflow"
author: "Andreas Brandmaier & Aaron Peikert"
date: "8/6/2019"
bibliography: Reproducible-Research.bib
output: html_document
---

@@

@@

The scientific community widely recognizes that these tools enhance reproducibility, and provides guidelines for their use (Literate Programming: @ruleTenSimpleRules2019; Version Control: @barbaHardRoadReproducibility2016; Dependency Management: @askrenUsingMakeReproducible2016; Containerization: @clyburne-sherinComputationalReproducibilityContainers2018). However, only a workflow using them in unison allows confidence in reproducing a scientific report. Within the community of users of the R Programming Language for each purpose, a specific software solution is the most popular. For literate programming, `rmarkdown`, for version control `git`, for dependency management `make` and for container `docker`. Each of them serves a meta-scientific valuable goal (reproducibility) and increases the productivity of the researcher. Each of these software solutions is exceptionally mighty, so to master them thoroughly often years of practice are required. Luckily the researcher can learn the minimal valuable subset needed to ensure reproducibility with little practice.

The authors believe that for sufficient chances of reproducibility, proper documentation is crucial since it answers the question "What steps have to be taken to reproduce the results?". The most precise documentation is computer code, and the most productive computer code is code that runs automatically without human interaction. These principles are what guides the use of `rmarkdown`, `git`, `make`, `docker` and their interaction.

@@

This folder, where everything resides that you need for analysis (code & data), is referred to as a Projekt. Working with Projects is exceptionally comfortable with RStudio an integrated development environment (IDE) for R. It is usefull to organize this Project in way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 @marwickPackagingDataAnalytical2018].

@@

Git & Github can do even more to help you to collaborate with your fellow researcher, but this is beyond the already broad scope of this tutorial. Another benefit of using Git & Github is that experimentation is highly encouraged since you can go back to any state quickly and even when you lose access to the file on your computer, everything is backed up on the Github servers.

@@

Note that this is the intended behavior and the result of a [bugfix](https://bugs.r-project.org/bugzilla/show_bug.cgi?id=17494) implemented from R 3.6.0 upwards.

@@

>>>>

---
title: "A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker"
shorttitle: "Reproducible Workflow"
author: "Aaron Peikert & Andreas M. Brandmaier"
date: "8/6/2019"
bibliography: Reproducible-Research.bib
output: html_document
---

@@

! TODO - Unterschied zwischen computational, empirical, and statistical reproducibility (cf. Stodden et al., 2014)

@@

The scientific community widely recognizes that these tools enhance reproducibility, and provides guidelines for their use (Literate Programming: @ruleTenSimpleRules2019; Version Control: @barbaHardRoadReproducibility2016; Dependency Management: @askrenUsingMakeReproducible2016; Containerization: @clyburne-sherinComputationalReproducibilityContainers2018). However, only a workflow using them in unison allows confidence in reproducing a scientific report. Within the community of users of the R Programming Language for each purpose, a specific software solution is the most popular. For literate programming, `rmarkdown`, for version control `git`, for dependency management `make` and for container `docker`. Each of them serves a meta-scientific valuable goal (reproducibility) and increases the productivity of the researcher. Each of these software solutions is highly flexible and powerful, so to master them thoroughly often years of practice are required. Luckily the researcher can learn the minimal valuable subset needed to ensure reproducibility with little practice.

We believe that for sufficient chances of reproducibility, proper documentation is crucial since it answers the question "What steps have to be taken to reproduce the results?". The most precise documentation is computer code, and the most productive computer code is code that runs automatically without human interaction. These principles are what guides the use of `rmarkdown`, `git`, `make`, `docker` and their interaction.

@@

Sometimes external requirements mean that the data cannot be stored and shared with the scripts. In most of the cases we have seen, these are either space constraints or privacy considerations. In these cases, of course, unrestricted reproducibility is not guaranteed. If this way of dividing data and scripts into different databases is unavoidable, we recommend that you crosscheck using checksums before loading data. To do this, a checksum must be created and filed at the time of the original analysis, and at the time of reproduction the current checksum must be compared with the filed checksum.

! TODO - Beispiel einbauen?

This folder, where everything resides that you need for analysis (code and data), is referred to as a Project. Working with Projects is exceptionally comfortable with RStudio an integrated development environment (IDE) for R. It is usefull to organize this Project in way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 @marwickPackagingDataAnalytical2018].

@@

Git and Github can do even more to help you to collaborate with your fellow researcher, such as formal means to track issues and their status (open/closed/resolved) and further means to managing and merging multiple, parallel versions of code (such as branches, pull requests, merges) but this is beyond the already broad scope of this tutorial. Another benefit of using Git and Github is that experimentation is highly encouraged since you can go back to any state quickly and even when you lose access to the file on your computer, everything is backed up on the Github servers.

@@

Note that this is the intended behavior and the result of a [bugfix](https://bugs.r-project.org/bugzilla/show_bug.cgi?id=17494) in the random number generator implemented from R 3.6.0 upwards.

@@

# Summary

! TODO - Summary of all this and all benefits

## Related approaches

! TODO

## Sharing reproducible workflows

How to best share a reproducible workflow? We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework (citation needed) or PsychData (citation needed). Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format. Typically, there is currently no single license that works for code, data, and media (such as text or figures). We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and moifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested in creating the workflow in the first place. In our experience, the Creative Commons license (CC-BY) is often appropriate for sharing texts, rmarkdown files, generated figures, and other media whereas scripts and any other computer code is often best shared according to the MIT license. Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.
